---
title: Flattened circadian glucocorticoid oscillations cause obesity due to increased
  lipid turnover and lipid uptake
author: Stefan Tholen, Kyle Kovary, Atefeh Rabiee, Ewa Bielczyk-Maczyńska, Wenting
  Yang, and Mary N. Teruel
date: 'Updated: `r format(Sys.time())`'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(sleuth)
library(readxl)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(eulerr)
```

### Introduction

This is a summary of the RNA-seq analysis with interfaces to explore the data collected for the project "Circadian glucocorticoid oscillations are required to maintain functional brown adipose tissue". This is an evolving project so expect the analysis, plot types, and interface to change over time. If anyone has suggestions please contact Kyle.


### Abstract

Chronic stressors flatten circadian glucocorticoid oscillations, which has been correlated with negative health outcomes including obesity. How flatter circadian glucocorticoid oscillations affect metabolism and fat storage remains unknown. Here we aimed to investigate consequences of flattened glucocorticoid oscillations in mice. We found that flattening glucocorticoid oscillations not only results in body weight gain mainly due to an increase in white fat depot mass, but also leads fat accumulation in brown adipose tissue and hyperinsulinemia. A transcriptomic analysis of white and brown adipose tissues revealed that flattened glucocorticoid oscillations resulted in a dysregulated lipid metabolism with a prominent role of Cd36. Indeed, Cd36 knockout mice are partially protected against body weight gain and lipid accumulation in the brown and visceral white fat depots induced by flattened glucocorticoid oscillations. These results provide insights how conditions associated with flattened glucocorticoid levels cause obesity.

### Sample Metadata Import

```{r metadata table, message=FALSE, warning=FALSE}
# Set up metadata table for WAT data
# In previous analysis (see WAT_Analysis.Rmd) samples 8,9,43 (mice 211, 223, 229)
# were determined to be outliers and will be excluded

wat_sample_list = read_xlsx(
  '/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq WAT analysis/Stefan Tholen_Samples_for_RNAseq_WAT.xlsx'
) %>%
  mutate(fat = 'WAT', lane = paste0(lane, '_WAT')) %>%
  dplyr::rename(sample_name = `Sample Name`) %>%
  unite('sample_name', c(sample_name, fat)) %>%
  dplyr::select(lane, mouse, sample_name)

path = list.dirs(
  '/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq WAT analysis/kallisto_output'
)[-c(1, 2)]
path = path %>% as_tibble() %>%
  separate(
    col = value,
    into = c('a', 'mouse'),
    sep = 'ST',
    remove = F
  ) %>% dplyr::select(value, mouse) %>%
  separate(
    col = mouse,
    into = c('b', 'mouse'),
    sep = '-',
    remove = T
  ) %>% dplyr::select(value, mouse)

s2c_wat = bind_cols(wat_sample_list, path[match(wat_sample_list$mouse, path$mouse), 1]) %>%
  dplyr::rename(path = value) %>%
  separate(
    sample_name,
    into = c('condition', 'delivery', 'day', 'replicate', 'depot'),
    remove = F
  ) %>%
  unite('treatment', c('condition', 'delivery'), remove = F) %>%
  dplyr::rename(sample = sample_name) %>%
  filter(!mouse %in% c(211, 223, 229)) #%>%
#unite('Treatment',c('Pellet','Delivery'), remove = F)

# Set up BAT metadata table
bat_sample_list = read_xlsx(
  '/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/Stefan Tholen_Samples_for_RNAseq_BAT.xlsx'
) %>%
  mutate(fat = 'BAT', lane = paste0(lane, '_BAT')) %>%
  dplyr::rename(sample_name = `Sample Name`) %>%
  unite('sample_name', c(sample_name, fat)) %>%
  dplyr::select(lane, mouse, sample_name)

path = list.dirs(
  '/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/kallisto_output/'
)[-c(1, 2)]
path = path %>% as_tibble() %>%
  separate(
    col = value,
    into = c('a', 'mouse'),
    sep = 'ST-',
    remove = F
  ) %>%
  dplyr::select(value, mouse)

s2c_bat = bind_cols(bat_sample_list, path[match(bat_sample_list$mouse, path$mouse), 1]) %>%
  dplyr::rename(path = value) %>%
  separate(
    sample_name,
    into = c('condition', 'delivery', 'day', 'replicate', 'depot'),
    remove = F
  ) %>%
  unite('treatment', c('condition', 'delivery'), remove = F) %>%
  dplyr::rename(sample = sample_name)

s2c <- rbind(s2c_wat, s2c_bat)
```

### Gene ID import

```{r import gene names, eval=TRUE, message=FALSE, warning=FALSE}

# Retrieve GeneNames
library(biomaRt)
mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
                         dataset = "mmusculus_gene_ensembl",
                         host = "dec2015.archive.ensembl.org")
# host = "ensembl.org")
t2g <- biomaRt::getBM(
  attributes = c(
    "ensembl_transcript_id",
    "transcript_version",
    "ensembl_gene_id",
    "external_gene_name",
    "description",
    "transcript_biotype",
    'external_transcript_name'
  ),
  mart = mart
)
t2g <- dplyr::rename(t2g,
                     target_id = ensembl_transcript_id,
                     ens_gene = ensembl_gene_id,
                     ext_gene = external_gene_name)

```

### Generate sleuth objects

```{r data import, eval=FALSE, message=FALSE, warning=FALSE}


so_all <- sleuth_prep(
  s2c,
  target_mapping = t2g,
  num_cores = max(1L, parallel::detectCores() - 1L),
  extra_bootstrap_summary = TRUE,
  read_bootstrap_tpm = TRUE
)
so_wat <- sleuth_prep(
  s2c_wat,
  target_mapping = t2g,
  num_cores = max(1L, parallel::detectCores() - 1L),
  extra_bootstrap_summary = TRUE,
  read_bootstrap_tpm = TRUE
)
so_bat <- sleuth_prep(
  s2c_bat,
  target_mapping = t2g,
  num_cores = max(1L, parallel::detectCores() - 1L),
  extra_bootstrap_summary = TRUE,
  read_bootstrap_tpm = TRUE
)

sleuth_save(so_all, '~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_all')
sleuth_save(so_wat, '~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_wat')
sleuth_save(so_bat, '~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_bat')

```

### Import sleuth objects

```{r message=FALSE, warning=FALSE}

so_all <- sleuth_load('~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_all')
so_wat <- sleuth_load('~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_wat')
so_bat <- sleuth_load('~/Documents/GitHub/ShinyApps/InVivoRNAseq/so_bat')

full_design <- model.matrix(formula( ~ treatment + depot + as.numeric(day)), data = s2c)

so_all <- sleuth_fit(so_all, formula = full_design, fit_name = "full")
so_all <- sleuth_fit(so_all, formula = ~ depot + as.numeric(day), fit_name = "reduced")
so_all <- sleuth_lrt(so_all, "reduced", "full")

#plot_qq(so, test = 'reduced:full', test_type = 'lrt', sig_level = 0.05)

lrt_results <- sleuth_results(so_all, 'reduced:full', test_type = 'lrt')
table(lrt_results[, "qval"] < 0.05)





full_design <- model.matrix(formula( ~ s2c_wat$treatment + s2c_wat$lane + s2c_wat$replicate, s2c_wat$day))

so_wat <- sleuth_fit(so_wat, formula = full_design, fit_name = "full")
so_wat <- sleuth_fit(so_wat, formula = ~ s2c_wat$lane + s2c_wat$replicate, fit_name = "reduced")
so_wat <- sleuth_lrt(so_wat, "reduced", "full")

lrt_results <- sleuth_results(so_wat, 'reduced:full', test_type = 'lrt')
table(lrt_results[, "qval"] < 0.05)

```


```{r}
hits = filter(lrt_results, qval < 0.05)
hits_dat = so_all$obs_norm %>% dplyr::select(target_id, sample, est_counts) %>% 
  filter(target_id %in% hits$target_id) %>% spread(key = sample, value = est_counts)
genes = hits_dat$target_id
genes = hits[match(genes,hits$target_id),'external_transcript_name']
hits_dat = as.matrix(hits_dat[,2:ncol(hits_dat)])
rownames(hits_dat) = genes

zeros = which(hits_dat <= 0, arr.ind = T)
for(i in 1:nrow(zeros)){
  hits_dat[zeros[i,1], zeros[i,2]] = 1
}

# Filter for conditions of interest
annotation = dplyr::select(s2c, sample, day, treatment, depot) %>% 
  #filter(Treatment %in% c('Control_na','Cort_injection', 'PBS_injection')) %>% 
  mutate(day = as.numeric(day)) %>% as.data.frame()

annotation = annotation[order(annotation$treatment),]
annotation_cols = list(ID = c(Control_na = '#bababa', Cort_injection = '#BF19C2', Cort_pellet = '#2CABCE',
                              Placebo_injection = '#2C9140', Placebo_pellet = '#AD2E24'))

cols <- colorRampPalette(brewer.pal(5, "Set1"))
mycolors <- cols(length(unique(annotation$Treatment)))
names(mycolors) <- unique(annotation$Treatment)
mycolors <- list(mycolors = mycolors)


hits_dat <- hits_dat[,colnames(hits_dat) %in% annotation$sample]
# Heatmap of Differentially Expressed Genes
hits_dat = t(apply(hits_dat, 1, function(x) x / mean(x)))


rownames(annotation) <- annotation$sample
annotation <- dplyr::select(annotation, -sample)

mypal = rev(colorRampPalette(brewer.pal(11, "RdBu"))(n=200))

pheatmap(log2(hits_dat), breaks = seq(-3,3,6/(length(mypal))), 
         color = mypal, 
         show_rownames = F,
         show_colnames = F,
         annotation_col = annotation,
         #annotation_colors = annotation_cols$ID,
         #cutree_cols = 2,
         clustering_distance_cols = 'correlation',
         border_color = F
         )
```


```{r}
pca = prcomp(t(hits_dat), center = T, scale. = T)

pcaSum = summary(pca)$importance %>% t() %>% as.data.frame() %>% mutate(PC = rownames(.)) %>%
  gather(key = plot, value = value, 1:3)
pcaSum$PC <- factor(pcaSum$PC, levels = paste0('PC',1:ncol(hits_dat)))

ggplot(pcaSum, aes(x = PC, y = value)) + geom_point() +
  facet_wrap(~plot, scales = 'free', ncol = 1) + theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(0, NA)


pcaDF = tibble(sample = rownames(pca$x), PC1 = pca$x[,1], PC2 = pca$x[,2], PC3 = pca$x[,3])
pcaDF = full_join(pcaDF, s2c, by = 'sample') %>% mutate(day = as.factor(day))
#pcaDF = pcaDF %>% mutate(outlier = ifelse(sample %in% c(8, 9, 43), T, F))

# 2D PCA between PC1 and PC2
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = treatment, size = day, shape = depot)) + geom_point(alpha = 0.5) + theme_bw()

ggplot(gather(pcaDF,key = PC, value = value, PC1:PC3), aes(x = value, fill = treatment)) + geom_dotplot(alpha = 0.5) + theme_bw() + facet_wrap(~PC, ncol = 1)

```


```{r}
s2cX <- s2c %>% filter(depot == 'WAT', treatment %in% c('Cort_pellet','Placebo_pellet'))
so_wat <- sleuth_prep(
  s2cX,
  target_mapping = t2g,
  aggregation_column = 'ext_gene',
  extra_bootstrap_summary = TRUE
)

## Transcript level comparison
# LRT w/ lane
so_wat <- sleuth_fit(so_wat, formula = ~ s2cX$treatment + s2cX$lane, fit_name = "full")
so_wat <- sleuth_fit(so_wat, formula = ~ s2cX$lane, fit_name = "reduced")

so_wat <- sleuth_lrt(so_wat, "reduced", "full")
lrt_ln_results <- sleuth_results(so_wat, 'reduced:full', test_type = 'lrt', pval_aggregate = FALSE) %>%
  rename(lrt_ln_qval = qval)
table(lrt_ln_results[, "lrt_ln_qval"] < 0.05)

# LRT w/o lane
so_wat <- sleuth_fit(so_wat, formula = ~ s2cX$treatment, fit_name = "full")
so_wat <- sleuth_fit(so_wat, formula = ~ 1, fit_name = "reduced")

so_wat <- sleuth_lrt(so_wat, "reduced", "full")
lrt_results <- sleuth_results(so_wat, 'reduced:full', test_type = 'lrt', pval_aggregate = FALSE) %>%
  rename(lrt_qval = qval)
table(lrt_results[, "lrt_qval"] < 0.05)

# Wald Test
models(so_wat)

so_wat <- sleuth_wt(so_wat, 's2cX$treatmentPlacebo_pellet')
wt_results <- sleuth_results(so_wat, 's2cX$treatmentPlacebo_pellet', test_type = 'wt', pval_aggregate = FALSE) %>%
  rename(wt_qval = qval)
table(wt_results[, "wt_qval"] < 0.05)

# Fold change / adjusted pval
so_wat_dat <- kallisto_table(so_wat) %>% group_by(target_id) %>%
  summarise(fc = mean(est_counts[treatment == 'Cort_pellet'], na.rm = T) / 
              mean(est_counts[treatment == 'Placebo_pellet'], na.rm = T),
            pval = t.test(est_counts[treatment == 'Cort_pellet'], 
                          est_counts[treatment == 'Placebo_pellet'])$p.value) %>%
  mutate(tt_qval = p.adjust(pval, 'BH'))
table(so_wat_dat[, "tt_qval"] < 0.05)

# Join 3 methods
joined_results <- dplyr::full_join(
  dplyr::select(lrt_results,
                target_id,
                ens_gene,
                ext_gene,
                lrt_qval),
  dplyr::select(so_wat_dat,
                target_id,
                fc,
                tt_qval),
  by = "target_id"
) %>%
  dplyr::full_join(
    dplyr::select(lrt_ln_results,
                  target_id,
                  lrt_ln_qval),
    by = 'target_id') %>%
  dplyr::full_join(
    dplyr::select(wt_results,
                  target_id,
                  b,
                  wt_qval),
    by = 'target_id'
  ) %>% as_tibble() %>% 
  dplyr::select(target_id, ens_gene, ext_gene, fc, b, everything())

# Compare models
# Num Hits
joined_results %>% dplyr::select(-ens_gene, -ext_gene, -fc, -b) %>% 
  rename(lrt = lrt_qval, lrt_ln = lrt_ln_qval, t.test = tt_qval, wald.test = wt_qval) %>%
  pivot_longer(names_to = 'test', values_to = 'qval', lrt:wald.test) %>%
  group_by(test) %>% summarise(num_hits = sum(qval < 0.05, na.rm = T)) %>%
  ggplot(., aes(x = fct_reorder(test, num_hits), y = num_hits)) + geom_bar(stat = 'identity') +
  ylab('Number of transcript hits') + xlab('Significance Test') + theme_bw() + theme(text = element_text(size=20))

# Venn Diagram
a = cbind(joined_results$tt_qval < 0.05,
          joined_results$wt_qval < 0.05,
          joined_results$lrt_qval < 0.05,
          joined_results$lrt_ln_qval < 0.05)
colnames(a) = c('t.test','wald.test','lrt','lrt.ln')
a[is.na(a)] = FALSE

plot(euler(a, shape = "ellipse"), quantities = TRUE)

####################
# GO Term analysis #
####################
library(clusterProfiler)
library(org.Mm.eg.db)
library(ggrepel)

# LRT w lane
gene.df <- clusterProfiler::bitr(dplyr::filter(joined_results, lrt_ln_qval < 0.05)$ext_gene, 
                                 fromType = 'SYMBOL', 
                                 toType = 'ENTREZID', 
                                 OrgDb = org.Mm.eg.db)

    
universe <- clusterProfiler::bitr(joined_results$ext_gene, 
                                  fromType = 'SYMBOL', 
                                  toType = 'ENTREZID', 
                                  OrgDb = org.Mm.eg.db)

    
goDF <- clusterProfiler::enrichGO(gene = gene.df$ENTREZID, 
                                  universe = universe$ENTREZID, 
                                  ont = 'ALL', 
                                  OrgDb = org.Mm.eg.db)

subsetFC <- joined_results %>% dplyr::rename(GeneName = ext_gene, FC = fc) %>% 
  mutate(FC = log2(FC))
    
goDF <- goDF %>% as_tibble() %>% group_by(ONTOLOGY) %>% mutate(rank = rank(p.adjust)) %>%
  rowwise() %>% dplyr::mutate(zscore = zScore(geneID, subsetFC, universe))

ggplot(filter(goDF, !ONTOLOGY == 'CC', p.adjust < 0.05), 
       aes(x = rank, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10) +
  xlim(0,20) + ylim(3,NA)

ggplot(filter(goDF, !ONTOLOGY == 'CC'), 
       aes(x = zscore, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10)

##########################
# LRT w/o lane
gene.df <- clusterProfiler::bitr(dplyr::filter(joined_results, lrt_qval < 0.05)$ext_gene, 
                                 fromType = 'SYMBOL', 
                                 toType = 'ENTREZID', 
                                 OrgDb = org.Mm.eg.db)

    
universe <- clusterProfiler::bitr(joined_results$ext_gene, 
                                  fromType = 'SYMBOL', 
                                  toType = 'ENTREZID', 
                                  OrgDb = org.Mm.eg.db)

    
goDF <- clusterProfiler::enrichGO(gene = gene.df$ENTREZID, 
                                  universe = universe$ENTREZID, 
                                  ont = 'ALL', 
                                  OrgDb = org.Mm.eg.db)

subsetFC <- joined_results %>% dplyr::rename(GeneName = ext_gene, FC = fc) %>% 
  mutate(FC = log2(FC))
    
goDF <- goDF %>% as_tibble() %>% group_by(ONTOLOGY) %>% mutate(rank = rank(p.adjust)) %>%
  rowwise() %>% dplyr::mutate(zscore = zScore(geneID, subsetFC, universe))

ggplot(filter(goDF, !ONTOLOGY == 'CC', p.adjust < 0.05), 
       aes(x = rank, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10) +
  xlim(0,20) + ylim(3,NA)

ggplot(filter(goDF, !ONTOLOGY == 'CC'), 
       aes(x = zscore, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10)

##########################
# WT
gene.df <- clusterProfiler::bitr(dplyr::filter(joined_results, wt_qval < 0.05)$ext_gene, 
                                 fromType = 'SYMBOL', 
                                 toType = 'ENTREZID', 
                                 OrgDb = org.Mm.eg.db)

    
universe <- clusterProfiler::bitr(joined_results$ext_gene, 
                                  fromType = 'SYMBOL', 
                                  toType = 'ENTREZID', 
                                  OrgDb = org.Mm.eg.db)

    
goDF <- clusterProfiler::enrichGO(gene = gene.df$ENTREZID, 
                                  universe = universe$ENTREZID, 
                                  ont = 'ALL', 
                                  OrgDb = org.Mm.eg.db)

subsetFC <- joined_results %>% dplyr::rename(GeneName = ext_gene, FC = fc) %>% 
  mutate(FC = log2(FC))
    
goDF <- goDF %>% as_tibble() %>% group_by(ONTOLOGY) %>% mutate(rank = rank(p.adjust)) %>%
  rowwise() %>% dplyr::mutate(zscore = zScore(geneID, subsetFC, universe))

ggplot(filter(goDF, !ONTOLOGY == 'CC', p.adjust < 0.05), 
       aes(x = rank, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10) +
  xlim(0,20) + ylim(3,NA)

ggplot(filter(goDF, !ONTOLOGY == 'CC'), 
       aes(x = zscore, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10)


##########################
# T-Test
gene.df <- clusterProfiler::bitr(dplyr::filter(joined_results, tt_qval < 0.05)$ext_gene, 
                                 fromType = 'SYMBOL', 
                                 toType = 'ENTREZID', 
                                 OrgDb = org.Mm.eg.db)

    
universe <- clusterProfiler::bitr(joined_results$ext_gene, 
                                  fromType = 'SYMBOL', 
                                  toType = 'ENTREZID', 
                                  OrgDb = org.Mm.eg.db)

    
goDF <- clusterProfiler::enrichGO(gene = gene.df$ENTREZID, 
                                  universe = universe$ENTREZID, 
                                  ont = 'ALL', 
                                  OrgDb = org.Mm.eg.db)

subsetFC <- joined_results %>% dplyr::rename(GeneName = ext_gene, FC = fc) %>% 
  mutate(FC = log2(FC))

goDF <- read_csv('~/Downloads/InVivoRNAseq    strongBAT RNA-seq Plotsstrong      strongWAT RNA-seq Plotsstrong (1).csv') %>% group_by(ONTOLOGY) %>% mutate(rank = rank(p.adjust))

goDF <- goDF %>% as_tibble() %>% group_by(ONTOLOGY) %>% mutate(rank = rank(p.adjust)) %>%
  rowwise() %>% dplyr::mutate(zscore = zScore(geneID, subsetFC, universe))

ggplot(filter(goDF, !ONTOLOGY == 'CC', p.adjust < 0.05), 
       aes(x = rank, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10) +
  xlim(0,20) + ylim(3,NA)

ggplot(filter(goDF, !ONTOLOGY == 'CC'), 
       aes(x = zscore, y = -log10(p.adjust), size = Count)) + 
  geom_point(alpha= 0.25) + facet_wrap(~ONTOLOGY, scales = 'free', ncol = 1) +
  geom_text_repel(data = filter(goDF, rank <= 20, !ONTOLOGY == 'CC'),
                    aes(label = Description), colour = 'black', size = 3, force = 10)

```


## Heatmap comparisons
```{r}
# Heatmap of intersect
hits = filter(so_wat_dat, target_id %in% hitIntersect)
hits_dat = so_wat$obs_norm %>% dplyr::select(target_id, sample, est_counts) %>% 
  spread(key = sample, value = est_counts)
genes = hits_dat$target_id
#genes = hits[match(genes,hits$target_id),'external_transcript_name']
hits_dat = as.matrix(hits_dat[,2:ncol(hits_dat)])
rownames(hits_dat) = genes

zeros = which(hits_dat <= 0, arr.ind = T)
for(i in 1:nrow(zeros)){
  hits_dat[zeros[i,1], zeros[i,2]] = 1
}

# Filter for conditions of interest
annotation = dplyr::select(s2cX, sample, day, treatment, depot) %>% 
  #filter(Treatment %in% c('Control_na','Cort_injection', 'PBS_injection')) %>% 
  mutate(day = as.numeric(day)) %>% as.data.frame()

annotation = annotation[order(annotation$treatment),]
annotation_cols = list(ID = c(Control_na = '#bababa', Cort_injection = '#BF19C2', Cort_pellet = '#2CABCE',
                              Placebo_injection = '#2C9140', Placebo_pellet = '#AD2E24'))

cols <- colorRampPalette(brewer.pal(5, "Set1"))
mycolors <- cols(length(unique(annotation$Treatment)))
names(mycolors) <- unique(annotation$Treatment)
mycolors <- list(mycolors = mycolors)


hits_dat <- hits_dat[,colnames(hits_dat) %in% annotation$sample]
# Heatmap of Differentially Expressed Genes
hits_dat = t(apply(hits_dat, 1, function(x) x / mean(x)))


rownames(annotation) <- annotation$sample
annotation <- dplyr::select(annotation, -sample)

mypal = rev(colorRampPalette(brewer.pal(11, "RdBu"))(n=200))

# Intersect
pheatmap(log2(hits_dat[rownames(hits_dat) %in% hitIntersect,]), 
         breaks = seq(-2,2,4/(length(mypal))), 
         color = mypal, 
         show_rownames = F,
         show_colnames = F,
         annotation_col = annotation,
         #annotation_colors = annotation_cols$ID,
         #cutree_cols = 2,
         clustering_distance_cols = 'correlation',
         border_color = F
         )

# LRT only
lrt_only <- uniqueLRT[(!uniqueLRT %in% uniqueFC)]
pheatmap(log2(hits_dat[rownames(hits_dat) %in% lrt_only,]), 
         breaks = seq(-2,2,4/(length(mypal))), 
         color = mypal, 
         show_rownames = F,
         show_colnames = F,
         annotation_col = annotation,
         #annotation_colors = annotation_cols$ID,
         #cutree_cols = 2,
         clustering_distance_cols = 'correlation',
         border_color = F
         )

# FC only
fc_only <- uniqueFC[(!uniqueFC %in% uniqueLRT)]
pheatmap(log2(hits_dat[rownames(hits_dat) %in% fc_only,]), 
         breaks = seq(-2,2,4/(length(mypal))), 
         color = mypal, 
         show_rownames = F,
         show_colnames = F,
         annotation_col = annotation,
         #annotation_colors = annotation_cols$ID,
         #cutree_cols = 2,
         clustering_distance_cols = 'correlation',
         border_color = F
         )
```


## TF Motifs WAT up vs down regulated genes
```{r}
# Export data for oPOSSUM analysis
joined_results %>% filter(lrt_ln_qval < 0.05, fc < 1/1.5) %>% 
  filter(!duplicated(ens_gene)) %>% 
  write_csv('~/Downloads/down_regulated.csv')

joined_results %>% filter(lrt_ln_qval < 0.05, fc > 2) %>% 
  filter(!duplicated(ens_gene)) %>% 
  write_csv('~/Downloads/up_regulated.csv')

# Load in oPOSSUM results
tf_up <- read_xlsx('~/Downloads/tf_motifs_up_regulated_wat.xlsx') %>% dplyr::select(TF, `Z-score`) %>%
  dplyr::rename(zscore_up = `Z-score`)

tf_down <- read_xlsx('~/Downloads/tf_motifs_down_regulated_wat.xlsx') %>% dplyr::select(TF, `Z-score`) %>%
  dplyr::rename(zscore_down = `Z-score`)

tf_all <- full_join(tf_up, tf_down, by = 'TF') %>% mutate(TF = toupper(TF))

ggplot(tf_all, aes(x = zscore_up, y = zscore_down)) + geom_point() + geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  geom_text_repel(data = filter(tf_all, zscore_up > 10, zscore_down < -10),
                    aes(label = TF), colour = 'black', size = 3, force = 10) +
  geom_text_repel(data = filter(tf_all, zscore_down > 10, zscore_up < -5),
                    aes(label = TF), colour = 'black', size = 3, force = 10)

tf_all <- joined_results %>% dplyr::select(ext_gene, fc, lrt_ln_qval) %>%
  dplyr::rename(TF = ext_gene) %>%
  mutate(direction = ifelse(lrt_ln_qval < 0.05 & fc > 1.5,
                            'Up',
                            ifelse(lrt_ln_qval < 0.05 & fc < 1/1.5,
                                   'Down',
                                   NA)),
         TF = toupper(TF)) %>%
  filter(!is.na(direction)) %>%
  right_join(tf_all, 'TF') %>% filter(!duplicated(TF))

ggplot(tf_all, aes(x = zscore_up, y = zscore_down, color = direction)) + 
  geom_vline(xintercept = 0, alpha = 0.5) + geom_hline(yintercept = 0, alpha = 0.5) + 
  geom_point(size = 4, alpha = 0.5, stroke = 0) + 
  geom_text_repel(data = filter(tf_all, zscore_up > 10, zscore_down < -5),
                    aes(label = TF, color = direction), size = 3, force = 10) +
  geom_text_repel(data = filter(tf_all, zscore_down > 10, zscore_up < -5),
                    aes(label = TF, color = direction), size = 3, force = 10) +
  theme_bw()

```

