---
title: "BAT RNA-seq Analysis"
author: "Stefan Tholen, Kyle M. Kovary, Mary N. Teruel"
date: "Updated: `r format(Sys.time())`"
output: html_document
---

### Introduction

This is a notebook to document the analysis of the RNA-seq data collected by Stefan from brown adipose tissue in mice that were treated with corticosterone pellet or injections to compare to sham pellet or injections over the course of 14 days. This is still in development, eventually I plan have the data analysis pipeline with quality control and analysis plots as well as the code used to analyze everything.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(sleuth)
library(DESeq2)
library(readxl)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
```

### Metadata Import

```{r metadata table, message=FALSE, warning=FALSE}
sample_list = read_xlsx('/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/Stefan Tholen_Samples_for_RNAseq_BAT.xlsx')

# Set up metadata table
path = list.dirs('/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/kallisto_output/')[-c(1,2)]
path = path %>% as.tibble() %>% 
  separate(col = value, into = c('a','mouse'), sep = 'ST-', remove = F) %>% 
  dplyr::select(value, mouse)

s2c = bind_cols(sample_list, path[match(sample_list$mouse, path$mouse),1]) %>%
  dplyr::rename(path = value, sample_name = `Sample Name`, Pellet = Treatment) %>% 
  unite('Treatment',c('Pellet','Delivery'), remove = F)

```

### Data Import

```{r data import, eval=FALSE, message=FALSE, warning=FALSE}

# Retrieve GeneNames
library(biomaRt)
tx2gene <- function(){
  mart <- biomaRt::useMart(biomart = "ensembl", dataset = "mmusculus_gene_ensembl")
  t2g <- biomaRt::getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id",
                                       "external_gene_name","entrezgene"), mart = mart)
  t2g <- dplyr::rename(t2g, target_id = ensembl_transcript_id,
                       ens_gene = ensembl_gene_id, ext_gene = external_gene_name)
  return(t2g)
}
t2g <- tx2gene()

so <- sleuth_prep(s2c, target_mapping = t2g, num_cores = max(1L, parallel::detectCores() - 1L), extra_bootstrap_summary = TRUE)

sleuth_save(so, '/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/so')

```

### DESeq2 Model

```{r message=FALSE, warning=FALSE}

so <- sleuth_load('/Users/kylekovary/Box Sync/kkovary/R Projects/RNAseq/Stefan RNA-seq BAT analysis/so')

full_design <- model.matrix(formula(~ s2c$Treatment + s2c$lane + s2c$Replicate, s2c$Day))

so <- sleuth_fit(so, full_design, 'full')
so <- sleuth_fit(so, formula = ~ s2c$lane + s2c$Replicate, fit_name = "reduced")
so <- sleuth_lrt(so, "reduced", "full")

#plot_qq(so, test = 'reduced:full', test_type = 'lrt', sig_level = 0.05)

lrt_results <- sleuth_results(so, 'reduced:full', test_type = 'lrt')
table(lrt_results[,"qval"] < 0.05)
```

### Select Differentially Expressed Genes

```{r message=FALSE, warning=FALSE}
hits = filter(lrt_results, qval < 0.05)
hits_dat = so$obs_norm_filt %>% dplyr::select(target_id, sample, est_counts) %>% filter(target_id %in% hits$target_id) %>% spread(key = sample, value = est_counts)
genes = hits_dat$target_id
genes = hits[match(genes,hits$target_id),'ext_gene']
hits_dat = as.matrix(hits_dat[,2:ncol(hits_dat)])
rownames(hits_dat) = genes

zeros = which(hits_dat <= 0, arr.ind = T)
for(i in 1:nrow(zeros)){
  hits_dat[zeros[i,1], zeros[i,2]] = 1
}
```

### Heatmap of Differentially Expressed Genes

```{r message=FALSE, warning=FALSE}
hits_dat = t(apply(hits_dat, 1, function(x) x / mean(x)))

annotation = dplyr::select(s2c, Day, Treatment) %>% as.data.frame()
rownames(annotation) <- s2c$sample

mypal = rev(colorRampPalette(brewer.pal(11, "RdBu"))(n=200))

pheatmap(log2(hits_dat), breaks = seq(-3,3,6/(length(mypal))), 
         color = mypal, 
         show_rownames = F,
         show_colnames = T,
         annotation_col = annotation,
         cutree_cols = 3)

```

There is a group of samples that cluster together for seemingly unknown reasons. We believe that these might be technical outliers. I will highlight these in the PCA plot to see if they are also outliers there. We might consider excluding these.

These samples are 28, 37, 31, 47, 39, 48, 50, 11, 24.

```{r message=FALSE, warning=FALSE, echo=TRUE}
outliers = filter(s2c, sample %in% c(28, 37, 31, 47, 39, 48, 50, 11, 24)) %>% 
  dplyr::select(sample, Pellet, Delivery, Day, Replicate) %>% 
  data.table::setorder(Delivery, Day, Replicate) %>% dplyr::rename(`Sample #` = sample, Condition = Pellet)

knitr::kable(outliers, row.names = F) %>% kable_styling()
```


### PCA of Differentially Expressed Genes

```{r message=FALSE, warning=FALSE}
pca = prcomp(t(hits_dat), center = T, scale. = T)

pcaSum = summary(pca)$importance %>% t() %>% as.data.frame() %>% mutate(PC = rownames(.)) %>%
  gather(key = plot, value = value, 1:3)
pcaSum$PC <- factor(pcaSum$PC, levels = paste0('PC',1:ncol(hits_dat)))

ggplot(pcaSum, aes(x = PC, y = value)) + geom_point() +
  facet_wrap(~plot, scales = 'free', ncol = 1) + theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(0, NA)


pcaDF = tibble(sample = as.numeric(rownames(pca$x)), PC1 = pca$x[,1], PC2 = pca$x[,2])
pcaDF = full_join(pcaDF, s2c, by = 'sample') %>% mutate(Day = as.factor(Day))
pcaDF = pcaDF %>% mutate(outlier = ifelse(sample %in% c(28, 37, 31, 47, 39, 48, 50, 11, 24), T, F))

# 2D PCA between PC1 and PC2
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = Pellet, size = Day, shape = Delivery)) + geom_point(alpha = 0.5) + theme_bw()

# 1D PCA plots since PC1 explains the majority of the variance
ggplot(gather(pcaDF,key = PC, value = value, PC1:PC2), aes(x = value, fill = Treatment)) + geom_dotplot(alpha = 0.5, binwidth = 5) + theme_bw() + facet_wrap(~PC, ncol = 1)

# Outlier colored PCA plots
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = outlier, size = Day, shape = Delivery)) + geom_point(alpha = 0.5) + theme_bw()

ggplot(gather(pcaDF,key = PC, value = value, PC1:PC2), aes(x = value, fill = outlier)) + geom_dotplot(alpha = 0.5, binwidth = 5) + theme_bw() + facet_wrap(~PC, ncol = 1)

```


### Enrched genes with and without outliers

In order to see if there is an impact on which genes are significantly differentially expressed with the outliers, we re-ran the DESeq model without them and looked for overlap between the gene sets.

```{r message=FALSE, warning=FALSE}

### MetaData w/o outliers ###

s2c2 = dplyr::filter(s2c, !sample %in% c(28, 37, 31, 47, 39, 48, 50, 11, 24))

### Retrieve GeneNames and import data ###
library(biomaRt)
tx2gene <- function(){
  mart <- biomaRt::useMart(biomart = "ensembl", dataset = "mmusculus_gene_ensembl")
  t2g <- biomaRt::getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id",
                                       "external_gene_name","entrezgene"), mart = mart)
  t2g <- dplyr::rename(t2g, target_id = ensembl_transcript_id,
                       ens_gene = ensembl_gene_id, ext_gene = external_gene_name)
  return(t2g)
}
t2g <- tx2gene()

so2 <- sleuth_prep(s2c2, target_mapping = t2g, num_cores = max(1L, parallel::detectCores() - 1L), extra_bootstrap_summary = TRUE)

### DESeq w/o outliers ###

full_design <- model.matrix(formula(~ s2c2$Treatment + s2c2$lane + s2c2$Replicate, s2c2$Day))

so2 <- sleuth_fit(so2, full_design, 'full')
so2 <- sleuth_fit(so2, formula = ~ s2c2$lane + s2c2$Replicate, fit_name = "reduced")
so2 <- sleuth_lrt(so2, "reduced", "full")

#plot_qq(so2, test = 'reduced:full', test_type = 'lrt', sig_level = 0.05)

lrt_results2 <- sleuth_results(so2, 'reduced:full', test_type = 'lrt')
table(lrt_results2[,"qval"] < 0.05)

### Select enriched genes ###

hits2 = filter(lrt_results2, qval < 0.05)
hits_dat2 = so2$obs_norm_filt %>% dplyr::select(target_id, sample, est_counts) %>% filter(target_id %in% hits2$target_id) %>% spread(key = sample, value = est_counts)
genes = hits_dat2$target_id
genes = hits2[match(genes,hits2$target_id),'ext_gene']
hits_dat2 = as.matrix(hits_dat2[,2:ncol(hits_dat2)])
rownames(hits_dat2) = genes

zeros = which(hits_dat2 <= 0, arr.ind = T)
for(i in 1:nrow(zeros)){
  hits_dat2[zeros[i,1], zeros[i,2]] = 1
}
```

### Overlap between the two hit lists

```{r message=FALSE, warning=FALSE}
hitIntersect = length(base::intersect(rownames(hits_dat),rownames(hits_dat2)))
uniqueFull = sum(!duplicated(rownames(hits_dat)))
uniqueRem = sum(!duplicated(rownames(hits_dat2)))

VennDiagram::draw.pairwise.venn(uniqueFull, uniqueRem, hitIntersect, category = c("All Samples", "Outliers Removed"), lty = rep("blank", 2), fill = c("light blue", "pink"), alpha = rep(0.5, 2))

#intersect = base::intersect(rownames(hits_dat),rownames(hits_dat2))
#fullExcl = rownames(hits_dat)[!duplicated(rownames(hits_dat))][!c(rownames(hits_dat)[!duplicated(rownames(hits_dat))] %in% intersect)]

#write_csv(as.tibble(intersect), '/Users/kylekovary/Downloads/intersect.csv')
#write_csv(as.tibble(fullExcl), '/Users/kylekovary/Downloads/fullExcl.csv')

write_csv(as.tibble(rownames(hits_dat)[!duplicated(rownames(hits_dat))]), '/Users/kylekovary/Downloads/withOutliers.csv')
write_csv(as.tibble(rownames(hits_dat2)[!duplicated(rownames(hits_dat2))]), '/Users/kylekovary/Downloads/noOutliers.csv')
write_csv(as.tibble(lrt_results$ext_gene[!duplicated(lrt_results$ext_gene)]), '/Users/kylekovary/Downloads/allGenes.csv')

```


### Overlap of hits: Combining time samples w/ FDR corrected p-values vs individual time points w/o FDR corrected p-values

```{r message=FALSE, warning=FALSE}


tpm = read_csv('../data/bat_normalized_data_genelevel_tpm.csv')

volDay3 = tpm %>% filter(Delivery == 'pellet', Day == 3) %>% group_by(GeneName) %>% 
  summarise(FC = mean(TPM[Treatment == 'Cort'], na.rm = T) / mean(TPM[Treatment == 'Sham'], na.rm = T),
            pval = t.test(TPM[Treatment == 'Cort'], TPM[Treatment == 'Sham'], na.rm = T)$p.value) %>%
  mutate(hit = ifelse(abs(log2(FC)) >= log2(1.5) &  pval <= 0.05, T, F)) %>%
  dplyr::select(GeneName, hit)


volDay7 = tpm %>% filter(Delivery == 'pellet', Day == 7) %>% group_by(GeneName) %>% 
  summarise(FC = mean(TPM[Treatment == 'Cort'], na.rm = T) / mean(TPM[Treatment == 'Sham'], na.rm = T),
            pval = t.test(TPM[Treatment == 'Cort'], TPM[Treatment == 'Sham'], na.rm = T)$p.value) %>%
  mutate(hit = ifelse(abs(log2(FC)) >= log2(1.5) &  pval <= 0.05, T, F)) %>%
  dplyr::select(GeneName, hit)


volDay14 = tpm %>% filter(Delivery == 'pellet', Day == 14) %>% group_by(GeneName) %>% 
  summarise(FC = mean(TPM[Treatment == 'Cort'], na.rm = T) / mean(TPM[Treatment == 'Sham'], na.rm = T),
            pval = t.test(TPM[Treatment == 'Cort'], TPM[Treatment == 'Sham'], na.rm = T)$p.value) %>%
  mutate(hit = ifelse(abs(log2(FC)) >= log2(1.5) &  pval <= 0.05, T, F)) %>%
  dplyr::select(GeneName, hit)


volPellet = tpm %>% filter(Delivery == 'pellet') %>% group_by(GeneName) %>% 
  summarise(FC = mean(TPM[Treatment == 'Cort'], na.rm = T) / mean(TPM[Treatment == 'Sham'], na.rm = T),
            pval = t.test(TPM[Treatment == 'Cort'], TPM[Treatment == 'Sham'], na.rm = T)$p.value) %>%
  mutate(pval = p.adjust(pval, 'BH'),
         hit = ifelse(abs(log2(FC)) >= log2(1.5) &  pval <= 0.05, T, F)) %>%
  dplyr::select(GeneName, hit)

a = cbind(volDay3$hit, volDay7$hit, volDay14$hit, volPellet$hit)
colnames(a) = c('Day3','Day7','Day14','AllFDR')

limma::vennDiagram(limma::vennCounts(a[,c(1,2)]), circle.col = c('blue','red','purple','green'))
limma::vennDiagram(limma::vennCounts(a[,c(1,3)]), circle.col = c('blue','red','purple','green'))
limma::vennDiagram(limma::vennCounts(a[,c(1,4)]), circle.col = c('blue','red','purple','green'))
limma::vennDiagram(limma::vennCounts(a[,c(2,3)]), circle.col = c('blue','red','purple','green'))
limma::vennDiagram(limma::vennCounts(a[,c(2,4)]), circle.col = c('blue','red','purple','green'))
limma::vennDiagram(limma::vennCounts(a[,c(3,4)]), circle.col = c('blue','red','purple','green'))

limma::vennDiagram(limma::vennCounts(a), circle.col = c('blue','red','purple','green'))
nHits = tibble(name = c('Day3','Day7','Day14','AllFDR'), 
               nHits = c(sum(a[,1]),sum(a[,2]),sum(a[,3]),sum(a[,4])))
ggplot(nHits, aes(x = name, y = nHits, fill = name)) + geom_bar(stat = 'identity') + theme_bw()

test = calculate.overlap(
  x = list(
    X = filter(volDay3, hit ==T),
    Y = filter(volDay7, hit ==T)
  )
)
venn.diagram(
   x = list(
    all = filter(volPellet, hit ==T)$GeneName,
    d3 = filter(volDay3, hit ==T)$GeneName
  ),
  filename = '3vsAll'
)
# venn.diagram(
# 	x = list(
# 		X = letters[1:10],
# 		Y = letters[5:20]
# 		#Z = letters[1:15]
# 		),
# 	filename = 'Venn',
# 	lwd = 4,
# 	#fill = c("cornflowerblue", "darkorchid1"),
# 	alpha = 0.75,
# 	#label.col = "white",
# 	cex = 4,
# 	fontfamily = "serif",
# 	fontface = "bold",
# 	#cat.col = c("cornflowerblue", "darkorchid1"),
# 	cat.cex = 3,
# 	cat.fontfamily = "serif",
# 	cat.fontface = "bold"
# 	#cat.dist = c(0.03, 0.03),
# 	#cat.pos = c(-20, 14)
# 	)
```


